Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-04-11 thirdPaper/draft/typeConeAssociahedra.tex
proofs of Lemmas 4.28 and 4.29

\begin{lemma}
\label{lemm::AR-conflations-and-sequences}
 Let~$\Lambda$ be any Artin algebra.  Let~$X\infl Y\defl Z$ be an almost-split conflation in~$K^{[-1,0]}(\proj \Lambda)$ such that~$Z$ is not in~$\add(\susp \Lambda)$.  Denote by~$F$ the functor~$\Hom{K^b}(\Lambda, -):K^{[-1,0]}(\proj \Lambda) \to \MOD \Lambda$.  Then there is an almost-split sequence
 \[
  0\to FX\to FY \to FZ \to 0
 \]
 in~$\MOD\Lambda$.
\end{lemma}
\begin{proof}

>>>>

\begin{lemma}
\label{lemm::AR-conflations-and-sequences}
 Let~$\Lambda$ be any Artin algebra.  Let~$X\overset{a}{\infl} Y \overset{b}{\defl} Z$ be an almost-split conflation in~$K^{[-1,0]}(\proj \Lambda)$ such that~$Z$ is not in~$\add(\susp \Lambda)$.  Denote by~$F$ the functor~$\Hom{K^b}(\Lambda, -):K^{[-1,0]}(\proj \Lambda) \to \MOD \Lambda$.  Then there is an almost-split sequence
 \[
  0\to FX\to FY \to FZ \to 0
 \]
 in~$\MOD\Lambda$.
\end{lemma}

\begin{proof}
The almost-split conflation induces an exact sequence
\[
F\susp^{-1}Y\to F\susp^{-1}Z \to FX \to FY\to FZ\to F\susp X.
\]
Because~$X$ is concentrated in non-positive degrees,~$FX=0$.
Let us verify that the map~$F\susp^{-1}Y\to F\susp^{-1}Z$ is surjective.
Let~$\Lambda \xrightarrow{f} \susp^{-1} Z$ be a morphism in~$K^b(\proj \Lambda)$.
Since~$Y\overset{b}{\to} Z$ is right almost-split, and~$Z$ is indecomposable not in~$\add(\susp\Lambda)$, the morphism~$\susp f$ factors through~$b$.
This implies the required surjectivity, and we have a short exact sequence
\[
 0 \to FX \xrightarrow{Fa} FY\xrightarrow{Fb} FZ\to 0.
\]
Because~$\MOD\Lambda$ and~$K^b(\proj \Lambda)$ are Krull--Schmidt, it is immediate to check that~$Fb$ (resp.~$Fa$) inherits from~$b$ (resp.~$a$) the property of being right (resp. left) almost-split.
\end{proof}

\begin{lemma}
\label{lemm::then-AR-seq-are-mutation}
Let~$\Lambda$ be an Artin algebra, let~$\tau X\infl E\defl X$ be an Auslander--Reiten conflation in~$K^{[-1,0]}(\proj\Lambda)$ and denote~$FX=M$.
Assume that:\yann{There might be some redundancies in the assumptions}
\begin{enumerate}
 \item $X\notin\add(\susp\Lambda)$;
 \item $M$ and~$\tau M$ are~$\tau$-rigid;
 \item $\Hom{\Lambda}(M,\tau^2 M)=0$;
 \item $\tau M$ is a brick.
\end{enumerate}
Then~$\tau X\infl E\defl X$ is a mutation conflation.
\end{lemma}

\begin{proof}
First note that the results~\cite[Lemmas 3.4 and 3.5]{AdachiIyamaReiten}, even though stated for finite dimensional algebras, hold with the same proofs for any Artin algebra.
For convenience of the reader, we recall those two results: Let~$Y, Z \in K^{[-1,0]}(\proj\Lambda)$ be indecomposable and not in~$\add(\susp\Lambda)$, and let~$P\in\proj\Lambda$. Then~$\Hom{K^b}(Y,\susp Z)=0$ if and only if~$\Hom{\Lambda}(FZ,\tau FY)=0$ and~$\Hom{K^b}(P,Y)=0$ if and only if~$\Hom{\Lambda}(P,FY)=0$.

Note also that Lemma~\ref{lemm::AR-conflations-and-sequences} implies that~$F\tau X$ is isomorphic to $\tau M$.

By assumption,~$X$ is not in~$\add(\susp\Lambda)$.
The same is true of~$\tau X$ since objects in~$\add(\susp\Lambda)$ are injective in~$K^{[-1,0]}(\proj\Lambda)$ and an Auslander--Reiten conflation does not split.
It thus follows from~\cite[Lemma 3.4 ]{AdachiIyamaReiten} that~$X$ and~$\tau X$ are rigid in $K^{[-1,0]}(\proj\Lambda)$.
We have to prove that both~$E\oplus X$ and~$E\oplus\tau X$ are rigid.

\vspace{7pt}
\noindent (1) $\Hom{K^b}(E,\susp X) = 0$.
\vspace{4pt}

The Auslander--Reiten conflation induces an exact sequence
\[
\Hom{K^b}(X,\susp X) \to \Hom{K^b}(E,\susp X) \to \Hom{K^b}(\tau X,\susp X).
\]
As already remarked,~$X$ is rigid so that the left-most term vanishes.
By assumption
\[
\Hom{\Lambda}(FX,\tau F\tau X) = \Hom{\Lambda}(M,\tau^2 M)=0,
\]
so that the right-most term also vanishes by~\cite[Lemma 3.4 ]{AdachiIyamaReiten}.

\vspace{7pt}
\noindent (2) $\Hom{K^b}(\tau X,\susp E) = 0$.
\vspace{4pt}

Since~$\Lambda$ is projective in~$K^{[-1,0]}(\proj\Lambda)$, the sequence~$\tau M \to FE \to M$ is short exact.
It induces a sequence
\[
\Hom{\Lambda}(M,\tau^2 M) \to \Hom{\Lambda}(FE,\tau^2 M) \to \Hom{\Lambda}(\tau M,\tau^2 M)
\]
from which we deduce that~$\Hom{\Lambda}(FE,\tau^2 M)=0$.
This implies~$\Hom{K^b}(\tau X,\susp E) = 0$ by~\cite[Lemma 3.4 ]{AdachiIyamaReiten}.

\vspace{7pt}
\noindent (3) $\Hom{K^b}(X,\susp E) = 0$.
\vspace{4pt}

Let us show that~$\Hom{\Lambda}(FE,\tau M)=0$.
Let~$FE \xrightarrow{f} \tau M$ be any morphism, and consider the short exact sequence~$0\to \tau M \xrightarrow{a} FE \xrightarrow{b} M \to 0$ induced by the Auslander--Reiten conflation.
Since~$\tau M$ is a brick, the composition~$fa$ is zero.
The morphism~$f$ then factors through the cokernel~$b$, which implies~$f=0$ since~$M$ is~$\tau$-rigid.

\vspace{7pt}
\noindent (4) $\Hom{K^b}(E,\susp E) = 0$.
\vspace{4pt}

The Auslander--Reiten conflation induces an exact sequence
\[
\Hom{K^b}(X,\susp E) \to \Hom{K^b}(E,\susp E) \to \Hom{K^b}(\tau X,\susp E).
\]
We can conclude by (2) and (3).

\vspace{7pt}
\noindent (5) $\Hom{K^b}(E,\susp \tau X) = 0$.
\vspace{4pt}

We have an exact sequence
\[
\Hom{K^b}(E,E) \to \Hom{K^b}(E,X) \to \Hom{K^b}(E,\susp \tau X) \to \Hom{K^b}(E,\susp E).
\]
The first map is surjective since~$E\to X$ is right almost split.
We can conclude by (4).
\end{proof}
